---
title: "Put a real title here"
subtitle: "SDS 192: MP5"
author: 
- name: Sethatevy Bong
  affiliation: Smith College
- name: Esa Schenck
  affiliation: Smith College
-name: Veronica Kessler  
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    code_folding: show
    df_print: paged
editor_options: 
  chunk_output_type: inline
references:
- id: xie2016bookdown
  type: book
  title: "Bookdown: Authoring books and technical documents with R markdown"
  author: 
  - family: Xie
    given: Yihui
  issued:
    year: 2016
  publisher: CRC Press
  url: https://bookdown.org/yihui/rmarkdown/
---


```{r setup, include=FALSE}
library(tidyverse)
library(sds192)
library(RMySQL)
library(leaflet)
library(sf)
db <- dbConnect(
  MySQL(), 
  host = "scidb.smith.edu", 
  user = "sds192", 
  password = "DSismfc@S", 
  dbname = "airlines"
)
knitr::opts_chunk$set(
  message = FALSE,
  echo = TRUE, 
  connection = db, 
  max.print = 20
)
```


## Introduction

What is this article about? [^github]

This article is about arguing whether age and gender are drivers for use the citibike system in New YorK and weather this could be a rationale tied to climate change perceptions. We also seek to expose the areas of higher confluence in the city for the use of these bikes.


[^github]: Please see [our GitHub repository](https://github.com/sds192-s20/mp3-group-*) for complete code. 

## Exploratory data analysis

```{sql, connection=db, output.var = "Trips_during_school_year"}
SELECT duration, birth_year, gender, start_station_id FROM citibike.trips
WHERE duration > 600
AND user_type = 'subscriber'
AND stop_time < '2017-06-01#' OR start_time > '2017-09-01#';
```


# We want to group those users younger than 30 and older than 30 to know which group uses the bike system more frequently

```{sql connection=db, output.var = "Trips"}
SELECT duration, birth_year, gender
 FROM citibike.trips
WHERE duration > 600
AND user_type = 'subscriber';

```

```{r}
Trips_Age <- Trips %>% 
  mutate(age_category = ifelse(birth_year >= 1987, "Thirthy or Younger", "Older Than Thirty"))
```


## In our first graph we plotted it the total trips people of each age group took in 2017. 

```{r}
Trips_Age %>% 
  group_by(birth_year, age_category) %>% 
   summarize(total = sum(n())/10000) %>%
  ggplot(aes(x = birth_year, y = total, fill= age_category)) + 
  geom_col()+
  xlim(c(1957, 2005))+
  labs(y= "Total Trips in Tens of Thousands")
```


### In our second graph we wanted to find out which gender is more likely to use the bike system in New York City. We found that men are more incline than women to ride the public bikes.

```{r}
Trips %>% 
  group_by(birth_year, gender) %>% 
  filter(birth_year >= 1957, gender != 0) %>%
  mutate(gender = ifelse(gender == 1, "Male", "Female")) %>% 
  summarize(total = sum(n())) %>%
  ggplot(aes(x = birth_year, y = total, fill = gender)) + 
  geom_col()
```



# Code Ben left as examples 
Make sure that you can get this chunk to run:


```{sql, connection=db}
SHOW DATABASES;
```

If you want to save the results of your query as a data frame, that you can use in any later R chunks, add the `connection` and `output.var` chunk options, like this:

```{sql, connection=db, output.var="first_ten_airports"}
SELECT * FROM airports LIMIT 10;
```

Now we have this object in R:

```{r}
first_ten_airports
```

Please see the [SQL chapter of the `bookdown` book](
https://rmarkdown.rstudio.com/authoring_knitr_engines.html%23sql#sql) for more information. 

You can cite references using parentheses or not. @xie2016bookdown wrote about this. See also [@xie2016bookdown].

```{r airport-map, fig.cap="Map of the first ten airports."}
library(leaflet)
leaflet() %>%
  addTiles() %>%
  addMarkers(
    data = first_ten_airports,
    popup = ~name
  )
```


```{sql connection=db, output.var = "Trips"}
SELECT duration, birth_year, gender
 FROM citibike.trips
WHERE duration > 600
AND user_type = 'subscriber';

```


# Station Data

```{sql connection=db, output.var = "months_station"}
SELECT station_id, name, num_starts, num_stops, avg_lat, avg_lon
 FROM citibike.station_months 
```

What stations get the most average rides (both starts and stops)?

```{r}
months_station_2 <- months_station %>% 
  #select(-station_id) %>% 
  mutate(trips = num_starts + num_stops) %>% 
  group_by(name) %>% 
  summarize(total_months = sum(n()),
          total_trips = sum(trips),
          avg_trips = total_trips/total_months,
          avg_lat = first(avg_lat),
          avg_lon = first(avg_lon)
         )
```


```{r}
months_station_2 %>% 
  ggplot(aes(x = name, y = avg_trips, fill = avg_trips)) +
  geom_col()
```
# Station Data

```{sql connection=db, output.var = "Popular_Stations"}
SELECT station_id, name, num_starts, num_stops, avg_lat, avg_lon
 FROM citibike.station_months 
 ORDER BY num_starts DESC;
```


```{r}
 Top_50_Stations <- Popular_Stations %>%
 head(50) 
station_pal <- colorNumeric(
  palette = "RdBu", 
  domain = Top_50_Stations %>%
    pull(num_starts) %>%
    range())

leaflet(Top_50_Stations) %>% 
  addTiles() %>%
  addCircles(lng = ~avg_lon, lat = ~avg_lat, weight = 3,
    radius = ~sqrt(num_stops)*3,
    color = ~station_pal(num_stops), popup = paste("<b>",Top_50_Stations$name, "</b>", "<br>",
                           "Number of Stops:", Top_50_Stations$num_stops), opacity = 50)
```






## Analysis

What did you discover? 

## Conclusion

What did you *learn* about the data set your explored?  

---

## Word count

```{r word_count, message=FALSE, echo=FALSE}
sds192::text_stats()
```


## Standards

In this assignment, we are attempting the following standards:

```{r buttons, echo=FALSE}
standard_button("markdown")
standard_button("aesthetics")
standard_button("wrangling")
standard_button("context")
standard_button("reshape")
standard_button("relational")
standard_button("ethics")
standard_button("github")
standard_button("iteration")
standard_button("function")
standard_button("spatial")
standard_button("leaflet")
standard_button("query")
```

## References

